---
output: github_document
---

```{r setup}
library(here)
library(tidyverse)
library(glue)
library(sf)
# devtools::install_github("or-house-vis/history")
library(history) 
```

# Seats held by two major parties

## Calculate seats over majority

Summarize seats held by each party each regular session:
```{r seats-by-party}
party_seats_held <- house_reps_regular %>% 
  group_by(session_year, party) %>% 
  count() %>% 
  ungroup() %>% 
  complete(session_year, party, fill = list(n = 0))
  # filled out so there is a row for every party in every year

party_seats_sum <- party_seats_held %>% 
  group_by(session_year) %>% 
  mutate(
    # put in explicit missings for the years with no records
    n = if_else(session_year %in% 1887:1891, NA_real_, n),
    filled_seats = sum(n),
    prop = n/filled_seats
  )

# check for record gap
party_seats_sum %>% 
  ggplot(aes(session_year, n)) +
    geom_line(aes(color = party)) +
    facet_wrap(~ party) +
    scale_color_manual(values = party_colors)
```

How many seats over majority?
```{r}
party_seats_over <- party_seats_sum %>% 
  mutate(
    required_thresh = floor(filled_seats/2),
    seats_over = if_else(n > required_thresh, n - required_thresh, 0)
  ) 
```

Just Republicans and Democrats
```{r}
dr_seats_over <- party_seats_over %>% 
  filter(party %in% c("Republican", "Democrat"))
```

Flip democrats to be negative to put them on LHS of zero
```{r}
dr_seats_over <- dr_seats_over %>% 
  mutate(
    side = if_else(party == "Democrat", -1, 1),
    seats_over_side = seats_over * side
  )
```

Static version of interactive timeline
```{r timeline}
dr_seats_over %>% 
  ggplot(aes(-session_year, seats_over_side)) + 
    geom_line(aes(color = party)) + 
    geom_area(aes(fill = party), alpha = 0.5, position = "identity") +
    coord_flip(clip = "off") +
    scale_x_continuous(expand = c(0, 0),
      breaks = c(-1860, seq(-1875, -2019, by = -25), -2019),
      labels = -1*c(-1860, seq(-1875, -2019, by = -25), -2019)) +
    scale_y_continuous("", limits = c(-30, 40),
      breaks = c(-30, -15, 0, 15, 30),
      labels = c(30, 15, 0, 15, 30), position = "right",
      sec.axis = dup_axis())  +
    scale_fill_manual(values = party_colors) +
    scale_color_manual(values = party_colors)
```

## Add captions

Caption for a session year, describing which of the two major parties has a majority and by how many seats:
```{r, warning = FALSE}
year_captions <- dr_seats_over %>% 
  arrange(session_year) %>% 
  group_by(session_year) %>% 
  filter(abs(seats_over_side) > 0) %>% 
  mutate(
    seats = abs(seats_over_side),
    color = party_colors[party],
    caption = glue('
      <strong>{session_year}</strong> <span style = "color:{color}">{party}s</span> have  {seats} seats over a majority.')
  ) %>% 
  select(session_year, caption) 
  
weird_year_captions <- tribble(
    ~ session_year, ~ caption,
    1874, "Neither party has a majority.",
    1887, "Party affiliations unknown.",
    1889, "Party affiliations unknown.",
    1891, "Party affiliations unknown.",
    2011, "Equally split between parties.",
    2012, "Equally split between parties.") %>% 
  mutate(caption = glue("<strong>{session_year}</strong> {caption}"))

year_captions <- bind_rows(year_captions, weird_year_captions)
```

Tabulations for all parties for each session year:
```{r, warning = FALSE}
# only need rows for parties represented in the session_year, 
# with parties listed from highest seats to lowest seats
party_seats_tab <- party_seats_sum %>% 
  filter(n != 0) %>% 
  arrange(session_year, desc(n)) %>% 
  select(session_year, party, n)

party_tables <- party_seats_tab %>% 
  nest() %>% 
  mutate(
    table = map(data, knitr::kable, 
      format = "html",
      row.names = FALSE,
      col.names = c("Party", "Seats")),
    table_str = as.character(table)
  ) %>% 
  select(session_year, table_str)
```

Put all together, will have captions duplicated for both Democrat and Republican rows:
```{r}
dr_seats_over_captioned <- dr_seats_over %>% 
  left_join(year_captions) %>% 
  left_join(party_tables) %>% 
  select(session_year, party, seats_over_side, caption, table_str) 
```

To Do: 

[ ] add way to test this with vega-lite that doesn't clobber published spec.

# Seats

Get geojson from https://github.com/or-house-vis/seat-layout
```{r}
url <- "https://raw.githubusercontent.com/or-house-vis/seat-layout/master/seats.json"
local <- here("docs", "data", "seats.json")
not_ok <- download.file(url, local)
stopifnot(!not_ok)

seats <- sf::read_sf(local)
seats %>% 
  select(col, row, seat)

# function to find seat number where
# seats are filled top to bottom, from outside in
find_seats <- function(ids, seats_df){
  seats_df %>% 
    arrange(row, abs(col - 4.5)) %>% 
    slice(seq_len(length(ids))) %>% 
    arrange(col, row) %>% 
    pull(seat)
}
```

# Individual Legislator data

## Get and assign to seats based on party and gender
```{r}
legislators <- house_reps_regular %>% 
  mutate(
    party = party %>% 
      as_factor() %>% 
      fct_explicit_na() %>% 
      fct_relevel("(Missing)", after = 2) %>% 
      fct_recode(`Unknown Affiliation` = "(Missing)")
    ) %>% 
  select(session_year, legislator, party, gender, residence, profession,
    notes) %>% 
  group_by(session_year) %>% 
  arrange(session_year, party, gender) %>% 
  mutate(id = 1:n(),
    seat = find_seats(id, seats))
```

Check on ordering:
```{r seat-assignment-check}
legislators %>% 
  filter(session_year == 2019) %>% 
  left_join(seats) %>% 
  ggplot() + 
    geom_sf(aes(fill =  party, alpha = gender)) +
    geom_sf_text(aes(label = seat))
```

## Legislator Captions

Clean up names and long strings in notes and residence:
```{r}
legislators_nice <- legislators %>% 
  mutate(name_split = str_split(legislator, ",", n = 2),
    # clean up names
    bad_name = map_lgl(name_split, ~ length(.) < 2),
    last_name = map_chr(name_split, 1) %>% str_trim(),
    first_name = ifelse(bad_name,
      list(c("")),
      map(name_split, 2)) %>% flatten_chr() %>% str_trim(),
    nice_name = str_c(first_name, last_name, sep = " "),
    # clean and wrap notes and residence
    notes = str_replace(notes, "\n", " ") %>% 
      str_wrap(width = 40) %>% 
      str_replace_all("\n", "<br/>"),
    residence = residence %>% 
      str_wrap(width = 40) %>% 
      str_replace_all("\n", "<br/>"))
```

Compose caption:
```{r, warning = FALSE}
legislators_captions <- legislators_nice  %>% 
  mutate(
    caption = glue('
      <strong>{nice_name}</strong><br/>{party}<br/>{gender}<br/>Representing {residence}.'
    ),
    caption = ifelse(is.na(notes), caption, 
      str_c(caption, "<br/><br/><em>", notes, "</em>"))
) %>% 
  select(session_year, legislator, caption) 
```

## Putting it all together

```{r}
people <- legislators %>% 
  left_join(legislators_captions) %>% 
  select(session_year, legislator, party, id, seat, caption) %>% 
  mutate(data = "legislators") 

years <- dr_seats_over_captioned %>% 
  mutate(data = "majority")

all <- years %>% 
  ungroup() %>% 
  bind_rows(people) %>% 
  mutate(session_year = ISOdate(year = session_year, month = 1, day = 1,
    tz = "utc")) 
```

Examine changes
```{r, eval = FALSE}
library(daff)
old <- read_csv(here("docs", "data", "vega-all.csv"))
diffs <- diff_data(all, old, ordered = FALSE)
render_diff(diffs)
```


Overwrite old data
```{r}
all %>%
  write_csv(here("docs", "data", "vega-all.csv"))
```

